---
title: "Stat380MiniProject3"
output: html_document
author: "Jeffery Yue and Ben Howland"
date: "04/28/25"
---

## Front Matter
```{r}
remove(list = ls())
library(tidyverse)
library(readxl)
library(glmnet)
library(rpart)
library(rattle)
COD <- read_excel("CODGames2_mp.xlsx")
```


## Problem 1
```{r}
COD2 = COD %>% 
  filter(FullPartial == "Full")
ggplot(data = COD2, mapping = aes(x = XPType, y = TotalXP))+geom_boxplot()+labs(x = "XP Type", y = "Total XP")
```
```{r}
summary = COD2 %>% group_by(XPType) %>% summarise(Min = min(TotalXP), Max = max(TotalXP), Mean = mean(TotalXP), Median = median(TotalXP), StdDev = sd(TotalXP))
print(summary)
```
The games with Double XP tend to have much higher total XP values than the games with just the 10% boost. The double XP games also have a higher standard deviation and more outliers indicating more variation. 


## Problem 2a

```{r}
#Code to determine if a player won that specific game
Splitresult = str_split_fixed(COD2$Result, "-", 2)
colnames(Splitresult) = c("Team", "Opponent")
Splitresult2 = data.frame(Splitresult)
Team = as.numeric(unlist(Splitresult2$Team))
Opponent = as.numeric(unlist(Splitresult2$Opponent))
Splitresult3 = data.frame(Team, Opponent)
codwins = Splitresult3 %>% mutate(Win = ifelse(Team > Opponent, TRUE, FALSE), Loss = ifelse(Team < Opponent, TRUE, FALSE), Draw = ifelse(Team == Opponent, TRUE, FALSE))

COD3 = COD2 %>% mutate(Win = codwins$Win)
```

```{r}
COD3 <- COD3 %>%
  filter(GameType == "HC - TDM")

XMat <- model.matrix(TotalXP ~ Eliminations + Deaths + Damage + XPType + Win, data = COD3)[,-1]
YVec <- COD3$TotalXP

set.seed(123)
lassoCV <- cv.glmnet(x = XMat, y = YVec,
                     family = "gaussian",
                     lambda = NULL,
                     standardize = TRUE,
                     nfolds = 10)
set.seed(NULL)

plot(lassoCV)

lassoCV$lambda.min
```

The optimal value for lambda is 107.5504 using lasso cross-validation with nfolds = 10.

```{r}
coefLamMin <- predict(lassoCV, s = lassoCV$lambda.min, type = "coefficients")
print(coefLamMin)
```

The equation for the lasso method is 
$$TotalXP = 1864.4863 + 438.7217x_{i, Eliminations} + 7368.3666x_{i, XPTypeDouble XP + 10%} + 1107.7902x_{i, WinTRUE}$$

```{r}
set.seed(123)
ridgeCV <- cv.glmnet(x = XMat, y = YVec,
                    family = "gaussian",
                    alpha = 0, 
                    lambda = NULL,
                    standardize = TRUE,
                    nfolds = 10)
set.seed(NULL)

plot(ridgeCV)

ridgeCV$lambda.min
```

The optimal value for lambda using Ridge with nfolds = 10 is 404.9202.

```{r}
coefRidMin <- predict(ridgeCV, s = ridgeCV$lambda.min, type = "coefficients")
print(coefRidMin)
```

The equation for the ridge technique is
$$TotalXP = 1696.2419 + 341.9336x_{i, Eliminations} + 13.1478x_{i, Deaths} + 3.4707x_{i, Damage} + 7114.0311x_{i, XPTypeDouble_XP_10%} + 1270.7964x_{i, WinTRUE}$$

The Lasso Method eliminated Deaths and Damage, while Ridge kept smaller values for them. They both had extremely large coefficients for XPType (double XP + 10%). This shows that that may be the most important variable for predicting the Total XP. Lasso made a simpler model, as it eliminated some variables.


## Problem 2b
```{r}
rpartobject = rpart(Score~TotalXP+Eliminations+Deaths+Damage+XPType+Win, method = "anova", data = COD3, minbucket = 15)
summary(rpartobject)
```
```{r}
fancyRpartPlot(rpartobject, cex = 0.7)
```
The 3 variables with the highest importance values are Damage(40), Eliminations(38), and TotalXP(10)



## Problem 2c

```{r}
XMatScaled <- scale(model.matrix(TotalXP ~ Eliminations + Deaths + Damage + XPType + Win, data = COD3)[,-1])

scaleddf <- data.frame(XMatScaled)

scaleddf$TotalXP <- COD3$TotalXP

lmRidge <- lm(TotalXP ~ ., data = scaleddf)
summary(lmRidge)
```

The estimated equation is $$TotalXP = 12590.3 + 6039.5x_{Eliminations} + 122.6x_{Deaths} - 3554.6x_{deaths} + 3818.5x_{XPType_DoubleXP_+_10%} + 689.9x_{WinTRUE}$$

```{r}
lmCoefs <- summary(lmRidge)$coefficients[-1, 1]
lmCoefs <- abs(lmCoefs)
print(lmCoefs)
```

The most important variables based on the magnitude of the coefficients of the liner regression model are Eliminations, with a coefficient of 6039.5451, XPType, with a coefficient of 3818.4585, and Damage, with a coefficient of 3554.6354. The Eliminations is the most important according to this model, followed by an XP Type of Double XP + 10%, followed by Damage. 
